On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-03-16 MS.tex
Fixed a ref Homogenized: "Fig."

Ultimately, data were recorded from 103 different articles \citep{Aiba:1997ho, Aiba:2009ft, Antunez:2001gk, Atkin:1998in, augspurger_light_1984, baltzer_physiological_2007, baraloto_differential_2006, baraloto_seed_2005, Bloor:2003fa,  Broncano:1998ck, Brown:1996ks, Bruhn:2000jz, Cai:2007ie, CamaraZapata:2003df, CastroDiez:1998gz, CastroDiez:2003jd, Cernusak:2008ds, Chao:2008hg, Chaturvedi:2011jx, Chaturvedi:2014ie, Chave:2009iy,  coomes_comparison_1998, coomes_greater_2009, Cornelissen:1996hf, Cornelissen:1997fd, Cornelissen:1998ta, Dalling:2004gs, DeBell:1994ix, Easdale:2009gv, fan_hydraulic_2012, Fayolle:2012bx, Fujimoto:2006fg, Galmes:2005fa, Gleason:2016, gleeson_plant_1994, grotkopp_toward_2002, Grubb:1996kr, Herault:2011dd, Hoffmann:2003ga, Huante:1995fd, Huante:1995tl, Huante:1998bk, HUNT:1997ge, Iida:2014ep, Iida:2014hq, king_influence_1994, king_tree_2005,  King:2006he, Kitajima:1994ez, Kohyama:2003dk, Kruger:2006fq, Lamers:2006dl, LopezIglesias:2014dk, Loveys:2002hj, Lusk:1997ga, Lusk:2002fma, Lusk:2013bq, Lusk:2013hz, MartinezVilalta:2010iq, McCormack:2012ch, Moya:2008ep, Muller-landau_interspecific_2004, Nascimento:2005kl, Osone:2008dr, Osunkoya:1994ip, Osunkoya:2010ef, Paz:2005kx, Poorter:1999fp, Poorter:2006vb, Poorter:2008iu, Poorter:2010co, Popma:1988tv, Prior:2004cv, Quero:2008jg, Read:2011du, Reich:1998ir, Reich:1998ja, Rossatto:2009gq, RuizRobleto:2005hc, Sack:2004hm, saldana-acosta_seedling_2009, SalgadoLuarte:2012ci, Saverimuttu:1996ih, Shen:2014fw, Shipley:2002fs, Stratton:2001ck, Thomas:1996do, Tomlinson:2014im, Veneklaas:2002dt, Villagra:2013do, Villar:2006cf, walters_are_1996, Walters:1993eu, Walters:1993hf, Wang:1998cs, Warren:2005bc, Westbrook:2011ja, Wright:1999ds, Wright:2000kw, Wright:2001gb, Wright:2003kb, Wright:2010tp, Zhang:1995bh}.

@@

To measure effect size, we used the cross-species Pearson product-moment correlation coefficient $r$ between growth rate and trait, weighted by the number of species contributing to the relationship. Results from studies that reported another coefficient -- such as coefficient of determination, Spearman's rho rank correlation, or Kendall's tau rank correlation -- were converted to correlation coefficients following \citet{Lajeunesse:2013tm}. As the magnitude of Pearson's correlation coefficient approaches $\pm$ 1.0, its distribution becomes skewed. Accordingly, we transformed the correlation $r$ into a metric $z$ with more-desirable statistical properties. The transformation was achieved via Fisher's z-transformation,
\[ z = \frac{1}{2} \ln \frac{1+r}{1-r},\]
which has a variance estimate of
\[ v_z= \frac{1}{n-3},\]
where $n$ is the number of species. We weighted effect size by the number of species used to calculate the correlation, instead of the raw number of individuals, because our questions primarily concerned the correlation between growth and traits across species and not within species. Throughout we report means and 95\% confidence intervals (CI) on effect size.

@@

For wood density and seed mass, the available results are consistent with our theoretical expectations, but it would be desirable to have substantially more data. Indeed, while it is possible to conduct a meta-analysis even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e highly uncertain estimates and CI often spanned zero). 
For wood density, results were consistent with the prediction for negative correlations with relative growth rate across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginally improved when stage was incorporated (p-value = 0.049). Note that the CIs did not cross zero for seedling and sapling stage, despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistent effect in the coefficient reported for each stage across studies. 
As expected, for seed mass the correlation with relative growth rate shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig \ref{fig:fig3}c, CIs overlapping zero). However, the correlations between seed mass and relative growth rate at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). We could not distinguish between these alternatives. But the number of coefficients reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficients reported rather than to a much smaller numbers of replicates.

For maximum height and Aarea, correlations with relative growth rate were also consistent with expectations, but our ability to test all hypotheses was at times limited due to low or no sample size. For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not significantly correlated at sapling stage (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage. While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.
For Aarea, results were consistent with a positive correlation with relative growth rate at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages, as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

Results from the entire dataset reinforced our conclusions for both SLA and wood density, and did not alter them for the others traits. For SLA and wood density, although the number of replicates doubled by using the entire dataset, effect sizes and overall results were similar to those reported above using the conservative criteria (Fig. \ref{fig:fig3}). For seed mass and Aarea, the additional replicates in the entire dataset allowed us to be more confident of the effect size for seedling stage (Fig \ref{fig:fig3}d, e). Comparison of Aarea effect between the conservative and relaxed criteria (Fig. \ref{fig:fig3}d), indicated little effect of stress (here mainly low light) on the correlation with relative growth rate at seedling stage as the correlation was still positive when 8 correlations measured under low light conditions were added. For maximum height, there were no additional replicates in the entire dataset; we still cannot conclude undoubtedly that the correlation is absent for smaller sizes.

@@

The only correlation that we hypothesized would change if absolute growth rate was used instead of relative growth rate, was for seed mass. Overall, the model fits did not improve when information about growth measurement (relative vs absolute growth rates) was incorporated in the model (see log likelihood ratio tests in Fig. \ref{fig:fig4}), thus supporting this claim. For all traits, effect sizes were non significantly different between relative growth rate and absolute growth rate when both measurements were available (e.g. CIs overlapping between relative and abolute growth rates for each stage, Fig. \ref{fig:fig4}). Effect sizes measured with a same number of observations for relative and absolute growth rate were generally very consistent (i.e. Hmax and wood density at adult stage in Fig. \ref{fig:fig4} b and c). However, for seed mass and wood density at seedling stage, the direction of effect sizes changed when correlations were measured with absolute versus relative growth rate. For seed mass, effect size at seedling stage with relative growth rate was negative and shifted to positive (but non-significant; CI included zero, Fig \ref{fig:fig4}d) with absolute growth rate. While these results were consistent with expectations, our ability to draw any firm conclusions was limited due to low sample size for correlation with absolute growth rate (i.e CIs overlapping, Fig. \ref{fig:fig4} d). For wood density, some differences were observed between studies reporting relative and absolute growth rates: in particular, the correlation at seedling stage shifted from negative with relative growth rate to positive (but non-significant; CI included zero, Fig. \ref{fig:fig4}b) with absolute growth rate. There may be several reasons for this difference, however, we note that the only available correlation coefficients recorded for absolute growth rate at seedling stage came from \citet{Augspurger:1984ct}, and unlike most other studies, they used wood density values estimated on adult plants together with growth data taken from seedlings. This suggests possible confounding between the type of growth measurement and origin of the trait data.

>>>>

Ultimately, data were recorded from 103 different articles \citep{Aiba:1997ho, Aiba:2009ft, Antunez:2001gk, Atkin:1998in, augspurger_light_1984, baltzer_physiological_2007, baraloto_differential_2006, baraloto_seed_2005, Bloor:2003fa,  Broncano:1998ck, Brown:1996ks, Bruhn:2000jz, Cai:2007ie, CamaraZapata:2003df, CastroDiez:1998gz, CastroDiez:2003jd, Cernusak:2008ds, Chao:2008hg, Chaturvedi:2011jx, Chaturvedi:2014ie, Chave:2009iy,  coomes_comparison_1998, coomes_greater_2009, Cornelissen:1996hf, Cornelissen:1997fd, Cornelissen:1998ta, Dalling:2004gs, DeBell:1994ix, Easdale:2009gv, fan_hydraulic_2012, Fayolle:2012bx, Fujimoto:2006fg, Galmes:2005fa, Gleason:2016, gleeson_plant_1994, grotkopp_toward_2002, Grubb:1996kr, Herault:2011dd, Hoffmann:2003ga, Huante:1995fd, Huante:1995tl, Huante:1998bk, HUNT:1997ge, Iida:2014ep, Iida:2014hq, king_influence_1994, king_tree_2005,  King:2006he, Kitajima:1994ez, Kohyama:2003dk, Kruger:2006fq, Lamers:2006dl, LopezIglesias:2014dk, Loveys:2002hj, Lusk:1997ga, Lusk:2002fma, Lusk:2013bq, Lusk:2013hz, MartinezVilalta:2010iq, McCormack:2012ch, Moya:2008ep, MullerLandau:2004dc, Nascimento:2005kl, Osone:2008dr, Osunkoya:1994ip, Osunkoya:2010ef, Paz:2005kx, Poorter:1999fp, Poorter:2006vb, Poorter:2008iu, Poorter:2010co, Popma:1988tv, Prior:2004cv, Quero:2008jg, Read:2011du, Reich:1998ir, Reich:1998ja, Rossatto:2009gq, RuizRobleto:2005hc, Sack:2004hm, saldana-acosta_seedling_2009, SalgadoLuarte:2012ci, Saverimuttu:1996ih, Shen:2014fw, Shipley:2002fs, Stratton:2001ck, Thomas:1996do, Tomlinson:2014im, Veneklaas:2002dt, Villagra:2013do, Villar:2006cf, walters_are_1996, Walters:1993eu, Walters:1993hf, Wang:1998cs, Warren:2005bc, Westbrook:2011ja, Wright:1999ds, Wright:2000kw, Wright:2001gb, Wright:2003kb, Wright:2010tp, Zhang:1995bh}.

@@

To measure effect size, we used the cross-species Pearson product-moment correlation coefficient $r$ between growth rate and trait, weighted by the number of species contributing to the relationship. Results from studies that reported another coefficient -- such as coefficient of determination, Spearman's rho rank correlation, or Kendall's tau rank correlation -- were converted to correlation coefficients following \citet{Lajeunesse:2013tm}. As the magnitude of Pearson's correlation coefficient approaches $\pm$ 1.0, its distribution becomes skewed. Accordingly, we transformed the correlation $r$ into a metric $z$ with more-desirable statistical properties. The transformation was achieved via Fisher's z-transformation,
\[ z = \frac{1}{2} \ln \frac{1+r}{1-r}\]
which has a variance estimate of
\[ v_z= \frac{1}{n-3}\]
where $n$ is the number of species. We weighted effect size by the number of species used to calculate the correlation, instead of the raw number of individuals, because our questions primarily concerned the correlation between growth and traits across species and not within species. Throughout we report means and 95\% confidence intervals (CI) on effect size.

@@

For wood density and seed mass, the available results are consistent with our theoretical expectations, but it would be desirable to have substantially more data. Indeed, while it is possible to conduct a meta-analysis even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e highly uncertain estimates and CI often spanned zero). 
For wood density, results were consistent with the prediction for negative correlations with relative growth rate across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginally improved when stage was incorporated (p-value = 0.049). Note that the CIs did not cross zero for seedling and sapling stage, despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistent effect in the coefficient reported for each stage across studies. 
As expected, for seed mass the correlation with relative growth rate shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig. \ref{fig:fig3}c, CIs overlapping zero). However, the correlations between seed mass and relative growth rate at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). We could not distinguish between these alternatives. But the number of coefficients reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficients reported rather than to a much smaller numbers of replicates.

For maximum height and Aarea, correlations with relative growth rate were also consistent with expectations, but our ability to test all hypotheses was at times limited due to low or no sample size. For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not significantly correlated at sapling stage (Fig. \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage. While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.
For Aarea, results were consistent with a positive correlation with relative growth rate at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages, as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

Results from the entire dataset reinforced our conclusions for both SLA and wood density, and did not alter them for the others traits. For SLA and wood density, although the number of replicates doubled by using the entire dataset, effect sizes and overall results were similar to those reported above using the conservative criteria (Fig. \ref{fig:fig3}). For seed mass and Aarea, the additional replicates in the entire dataset allowed us to be more confident of the effect size for seedling stage (Fig. \ref{fig:fig3}d and e). Comparison of Aarea effect between the conservative and relaxed criteria (Fig. \ref{fig:fig3}d), indicated little effect of stress (here mainly low light) on the correlation with relative growth rate at seedling stage as the correlation was still positive when 8 correlations measured under low light conditions were added. For maximum height, there were no additional replicates in the entire dataset; we still cannot conclude undoubtedly that the correlation is absent for smaller sizes.

@@

The only correlation that we hypothesized would change if absolute growth rate was used instead of relative growth rate, was for seed mass. Overall, the model fits did not improve when information about growth measurement (relative vs absolute growth rates) was incorporated in the model (see log likelihood ratio tests in Fig. \ref{fig:fig4}), thus supporting this claim. For all traits, effect sizes were non significantly different between relative growth rate and absolute growth rate when both measurements were available (e.g. CIs overlapping between relative and abolute growth rates for each stage, Fig. \ref{fig:fig4}). Effect sizes measured with a same number of observations for relative and absolute growth rate were generally very consistent (i.e. Hmax and wood density at adult stage in Fig. \ref{fig:fig4}b and c). However, for seed mass and wood density at seedling stage, the direction of effect sizes changed when correlations were measured with absolute versus relative growth rate. For seed mass, effect size at seedling stage with relative growth rate was negative and shifted to positive (but non-significant; CI included zero, Fig. \ref{fig:fig4}d) with absolute growth rate. While these results were consistent with expectations, our ability to draw any firm conclusions was limited due to low sample size for correlation with absolute growth rate (i.e CIs overlapping, Fig. \ref{fig:fig4}d). For wood density, some differences were observed between studies reporting relative and absolute growth rates: in particular, the correlation at seedling stage shifted from negative with relative growth rate to positive (but non-significant; CI included zero, Fig. \ref{fig:fig4}b) with absolute growth rate. There may be several reasons for this difference, however, we note that the only available correlation coefficients recorded for absolute growth rate at seedling stage came from \citet{Augspurger:1984ct}, and unlike most other studies, they used wood density values estimated on adult plants together with growth data taken from seedlings. This suggests possible confounding between the type of growth measurement and origin of the trait data.
